---
title: "PRS prediction Tin et al"
author: "Ruth Topless"
date: "`r format(Sys.Date())`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(DT)
```

Questions:
Which Gout affection status do I use? visit 0
What age do I adjust by? age at visit 0

Effect sizes all converted to risk  
Male and Female analaysis uses the effect sizes from the combined GWAS to weight the SNPs. not sex specific weighting. 


114 SNPs

```{r}
#GRS pheWAS in ukbiobank

library(tidyverse)
pheno<-read.delim("/Volumes/staff_groups/merrimanlab/Merriman_Documents/Ruth/UKBioBank/Phenotypes/data/inital_pheno_visit0_decoded.txt", stringsAsFactors = F)
pheno2<-pheno %>% select(f_eid, f_31_0_0, f_34_0_0,f_21000_0_0, f_189_0_0, f_54_0_0, f_21001_0_0, f_20116_0_0, f_21003_0_0)
colnames(pheno2)<-c("f_eid", "SEX", "YEAR","ETHNICITY", "TOWNSEND", "SITE", "BMI", "SMOKER", "AGE")
urate<-read.delim("/Volumes/staff_groups/merrimanlab/Merriman_Documents/Ruth/UKBioBank/Phenotypes/data/ukb_urate_umolL_v0.txt", stringsAsFactors = F)
colnames(urate)<-c("f_eid", "URATE")
pheno2<-merge(pheno2, urate, by="f_eid", all.x= T)

affstat<-read.delim("/Volumes/archive/merrimanlab/raid_backup/UKbiobank/genetic_files/gout_gwas_covar.20210201.covar")
pheno2<-merge(pheno2,affstat[,c("IID","plink_goutaff")], by.x="f_eid",by.y="IID")
illness <-read.delim("/Volumes/staff_groups/merrimanlab/Merriman_Documents/Ruth/UKBioBank/Phenotypes/data/illness_v0_upload_2019.txt")
srgout0<-illness %>% dplyr::filter(str_detect(GOUT_SR_0, "1")) %>% pull(f_eid) %>% unique()

pheno2$gout_V0=NA
pheno2$gout_V0[which(pheno2$f_eid %in% srgout0)]=1
pheno2$gout_V0[is.na(pheno2$gout_V0)]=0
pheno2$gout_V0[which(pheno2$plink_goutaff==2 & pheno2$gout_V0==0)]=NA

pheno2$hyperu<- NA
pheno2$hyperu[which(pheno2$URATE>=420)]=0
pheno2$hyperu[which(pheno2$gout_V0==1)]=1

goutprs<-read.delim("/Volumes/staff_groups/merrimanlab/Merriman_Documents/Ruth/2021_projects/PRS/UKBB_tin114_grs_2021_for_PRS.txt")
colnames(goutprs)<-c("f_eid","GoutPRS")
pheno2<-merge(pheno2,goutprs, by="f_eid", all.x = T)

pheno2$ETHNICCLASS<-pheno2$ETHNICITY
pheno2$ETHNICCLASS <- plyr::mapvalues(pheno2$ETHNICCLASS, from = c("African","White and Black African", "Black or Black British","Caribbean", "Any other Black background","White and Black Caribbean", "Asian or Asian British","Chinese","Indian","Pakistani","White and Asian","Any other Asian background", "Bangladeshi", "British","Any other white background", "Irish","White", "Prefer not to answer","Any other mixed background","Do not know","Mixed","Other ethnic group"), to = c("BLACK","BLACK","BLACK","BLACK","BLACK","BLACK","ASIAN","ASIAN","ASIAN","ASIAN","ASIAN","ASIAN","ASIAN","WHITE", "WHITE","WHITE","WHITE", "OTHER","OTHER","OTHER","OTHER","OTHER"))
pheno2<-pheno2[which(pheno2$ETHNICCLASS=="WHITE"),]
pheno2<-pheno2[is.na(pheno2$GoutPRS)==F,]

from = range(pheno2$GoutPRS, na.rm = T)[1] 
to = range(pheno2$GoutPRS, na.rm = T)[2] 
by = (to-from)/10
breaksgout <- seq( from = from, to = to, by = by )
# bucketing values into bins
pheno2<-pheno2 %>% mutate(GoutPRS_bins = cut(pheno2$GoutPRS, breaks=breaksgout, include.lowest=TRUE, right=FALSE))

```

## Part A Gout PRS bins

```{r}

sumfilegout<-as.data.frame(table(pheno2$GoutPRS_bins))
sumfilegout$controls<-table(pheno2$GoutPRS_bins, pheno2$gout_V0)[1:10]
sumfilegout$gout<-table(pheno2$GoutPRS_bins, pheno2$gout_V0)[11:20]
sumfilegout$prevalencegout<- round(x = (sumfilegout$gout/sumfilegout$Freq)*100, digits = 3)

sumfilegout %>%  datatable(., caption = "GoutPRS bins and gout prevalence", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)

f<-ggplot(sumfilegout, aes(Var1,Freq))+geom_col()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20)))
f
a<-sumfilegout %>% mutate(Var1_number = as.numeric(Var1)) %>% ggplot( aes(Var1_number,prevalencegout))+geom_line()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + scale_x_continuous(labels = as.character(sumfilegout$Var1), breaks = 1:nlevels(sumfilegout$Var1))
a
```


## Part B Gout PRS bins
```{r}
results= function(test, i=2){
  matrixoutput = matrix(nrow=1, ncol=6)
  matrixoutput[1] <- names(coefficients(test)[i])
  matrixoutput[2] <- as.numeric(exp(coefficients(test))[i]) #OR
  matrixoutput[3] <- as.numeric(exp(coef(summary(test))[i, 1] - 1.96 * coef(summary(test))[i, 2])) #LCI
  matrixoutput[4] <- as.numeric(exp(coef(summary(test))[i, 1] + 1.96 * coef(summary(test))[i, 2])) #UCI
  matrixoutput[5] <- as.numeric(coef(summary(test))[i,c(4)]) #p value
  matrixoutput[6] <- test$df.null+1
  return ((matrixoutput))
}

resulttable <- as.data.frame(matrix(ncol=6, nrow=0))
colnames(resulttable)<-c("name","OR","LCI","UCI", "P","N" )
pheno2$GoutPRS_bins<-factor(pheno2$GoutPRS_bins, levels =c("[4.73,5.02)","[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)",  "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]" ))
test<-glm(gout_V0~GoutPRS_bins+AGE+SEX,data=pheno2) 

for(b in seq(2:13)){
    resulttable[b,]<-results(test=test,i=b)
}
resulttable[1,]<-c("GoutPRS_bins[4.73,5.02)","1","1","1",NA, NA)
resulttable[,1]<-c("[4.73,5.02)","[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)",  "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]", "Age", "Sex")
resulttable$OR<-as.numeric(resulttable$OR)
resulttable$UCI<-as.numeric(resulttable$UCI)
resulttable$LCI<-as.numeric(resulttable$LCI)
resulttable$P<-as.numeric(resulttable$P)

resulttable$sig<-NA
resulttable$sig[which(resulttable$P<0.05)]="*"
resulttable$sig[which(resulttable$P<0.01)]="**"
resulttable$sig[which(resulttable$P<0.001)]="***"

resulttable$name <- factor(resulttable$name, levels = c("[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[4.73,5.02)", "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]", "Age", "Sex"))

library(ggrepel)
ggplot(data=resulttable, aes(x=name, y=OR)) +
  geom_bar(stat="identity", color="black") +
  labs(title = "All", y= "Odds Ratio (Log scale)", x = "Risk Factor") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + 
  scale_y_log10() + 
  geom_errorbar(aes(ymin=LCI, ymax=UCI), width=.2, position=position_dodge(.9)) +
  geom_text_repel(label = resulttable$sig, nudge_y = 0.015, aes(segment.alpha=0))

#fix significance
#library(svglite)
#ggsave("~/Desktop/GoutCohort_barchart.svg")

resulttable %>%  datatable(., caption = "Logisitic regression Gout ~ GRS bins ", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)


```


## REPEAT tests in hyperU subset
new bins made from reduced dataset
```{r}

HyperU<-pheno2[is.na(pheno2$hyperu)==F,]
breaksgout2 <- seq(from = range(HyperU$GoutPRS, na.rm = T)[1], to = range(HyperU$GoutPRS, na.rm = T)[2], by = (range(HyperU$GoutPRS, na.rm = T)[2] - range(HyperU$GoutPRS, na.rm = T)[1])/10)
# bucketing values into bins
HyperU<-HyperU %>% mutate(GoutPRS_bins = cut(HyperU$GoutPRS, breaks=breaksgout2, include.lowest=TRUE, right=FALSE))

sumfilegout_HU<-as.data.frame(table(HyperU$GoutPRS_bins))
sumfilegout_HU$controls<-table(HyperU$GoutPRS_bins, HyperU$gout_V0)[1:10]
sumfilegout_HU$gout<-table(HyperU$GoutPRS_bins, HyperU$gout_V0)[11:20]
sumfilegout_HU$prevalencegout<- round(x = (sumfilegout_HU$gout/sumfilegout_HU$Freq)*100, digits = 3)

sumfilegout_HU %>%  datatable(., caption = "GoutPRS bins and gout prevalence in HyperU", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)
HyperU$GoutPRS_bins<-factor(HyperU$GoutPRS_bins, levels =c("[4.83,5.1)", "[3.77,4.03)",  "[4.03,4.3)",  "[4.3,4.57)", "[4.57,4.83)",   "[5.1,5.37)", "[5.37,5.63)",  "[5.63,5.9)", "[5.9,6.17)", "[6.17,6.43]"))
test<-glm(gout_V0~GoutPRS_bins+AGE+SEX,data=HyperU) 
resulttable_HU <- as.data.frame(matrix(ncol=6, nrow=0))
colnames(resulttable_HU)<-c("name","OR","LCI","UCI", "P","N" )

for(b in seq(2:13)){
    resulttable_HU[b,]<-results(test=test,i=b)
}
resulttable_HU[1,]<-c("[4.83,5.1)","1","1","1",NA, NA)
resulttable_HU[,1]<-c("[4.83,5.1)", "[3.77,4.03)",  "[4.03,4.3)",  "[4.3,4.57)", "[4.57,4.83)", "[5.1,5.37)", "[5.37,5.63)",  "[5.63,5.9)", "[5.9,6.17)", "[6.17,6.43]", "Age", "Sex")
resulttable_HU$OR<-as.numeric(resulttable_HU$OR)
resulttable_HU$UCI<-as.numeric(resulttable_HU$UCI)
resulttable_HU$LCI<-as.numeric(resulttable_HU$LCI)
resulttable_HU$P<-as.numeric(resulttable_HU$P)

resulttable_HU$sig<-NA
resulttable_HU$sig[which(resulttable_HU$P<0.05)]="*"
resulttable_HU$sig[which(resulttable_HU$P<0.01)]="**"
resulttable_HU$sig[which(resulttable_HU$P<0.001)]="***"

resulttable_HU$name <- factor(resulttable_HU$name, levels = c( "[3.77,4.03)",  "[4.03,4.3)",  "[4.3,4.57)", "[4.57,4.83)",  "[4.83,5.1)", "[5.1,5.37)", "[5.37,5.63)",  "[5.63,5.9)", "[5.9,6.17)", "[6.17,6.43]", "Age", "Sex"))

library(ggrepel)
ggplot(data=resulttable_HU, aes(x=name, y=OR)) +
  geom_bar(stat="identity", color="black") +
  labs(title = "Tim_GRS in UKBB_HyperU", y= "Odds Ratio (Log scale)", x = "Risk Factor") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + 
  scale_y_log10() + 
  geom_errorbar(aes(ymin=LCI, ymax=UCI), width=.2, position=position_dodge(.9)) +
  geom_text_repel(label = resulttable_HU$sig, nudge_y = 0.015, aes(segment.alpha=0))

resulttable_HU %>%  datatable(., caption = "Logisitic regression Gout ~ GRS bins in HyperU", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)

```



## Part C for Gout GRS

```{r}

#Below are various packages I have tried for calculating the AUROC and AUPRC using the risk score. However these get different results to the ones I get in the next section using a prediction model rather than just the score. I don't exactly know what is going on here. 

test_dataset<-pheno2[,c("f_eid","SEX","AGE","GoutPRS","gout_V0")]
test_dataset<-na.omit(test_dataset)
colnames(test_dataset)[5]="affstat"

library(pROC) # install with install.packages("pROC")
library(PRROC) # install with install.packages("PRROC")
library(ROCR) #install with install.packages("ROCR")

pROC::roc(response = test_dataset$affstat, test_dataset$GoutPRS, plot=TRUE, legacy.axes=TRUE, percent=TRUE, xlab="False Positive Percentage", ylab="True Postive Percentage")

roc<-PRROC::roc.curve(scores.class0 = test_dataset[test_dataset$affstat=="1",]$GoutPRS, scores.class1 = test_dataset[test_dataset$affstat=="0",]$GoutPRS, curve = TRUE)
roc
plot(roc)

pr<-pr.curve(scores.class0 = test_dataset[test_dataset$affstat=="1",]$GoutPRS, scores.class1 = test_dataset[test_dataset$affstat=="0",]$GoutPRS, curve = TRUE)
pr
plot(pr)

```


# Part C for AGE and SEX GRS

Tin et al. using smaller group of UKBB Europeans got 0.8 AUROC for age and sex prediction of gout. Much higher than this. 

```{r}

pROC::roc(response = test_dataset$affstat, test_dataset$AGE+test_dataset$SEX, plot=TRUE, legacy.axes=TRUE, percent=TRUE, xlab="False Positive Percentage", ylab="True Postive Percentage")

roc<-PRROC::roc.curve(scores.class0 = test_dataset[test_dataset$affstat=="1",]$AGE+test_dataset[test_dataset$affstat=="1",]$SEX, scores.class1 = test_dataset[test_dataset$affstat=="0",]$AGE+test_dataset[test_dataset$affstat=="0",]$SEX, curve = TRUE)
roc
plot(roc)

pr<-pr.curve(scores.class0 = test_dataset[test_dataset$affstat=="1",]$AGE+test_dataset[test_dataset$affstat=="1",]$SEX, scores.class1 = test_dataset[test_dataset$affstat=="0",]$AGE+test_dataset[test_dataset$affstat=="0",]$SEX, curve = TRUE)
pr
plot(pr)

```

# males

```{r}
males<-pheno2[which(pheno2$SEX==1),]
# breaksgout2 <- seq(from = range(males$GoutPRS, na.rm = T)[1], to = range(males$GoutPRS, na.rm = T)[2], by = (abs(range(males$GoutPRS, na.rm = T)[1]) + range(males$GoutPRS, na.rm = T)[2])/10)
# # bucketing values into bins
# males<-males %>% mutate(GoutPRS_bins = cut(males$GoutPRS, breaks=breaksgout2, include.lowest=TRUE, right=FALSE))
males$GoutPRS_bins <- factor(males$GoutPRS_bins, levels = c("[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[4.73,5.02)", "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]"))
sumfilegout_males<-as.data.frame(table(males$GoutPRS_bins))
sumfilegout_males$controls<-table(males$GoutPRS_bins, males$gout_V0)[1:10]
sumfilegout_males$gout<-table(males$GoutPRS_bins, males$gout_V0)[11:20]
sumfilegout_males$prevalencegout<- round(x = (sumfilegout_males$gout/sumfilegout_males$Freq)*100, digits = 3)

f<-ggplot(sumfilegout_males, aes(Var1,Freq))+geom_col()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20)))
f
a<-sumfilegout_males %>% mutate(Var1_number = as.numeric(Var1)) %>% ggplot( aes(Var1_number,prevalencegout))+geom_line()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + scale_x_continuous(labels = as.character(sumfilegout$Var1), breaks = 1:nlevels(sumfilegout$Var1))
a

sumfilegout_males %>%  datatable(., caption = "GoutPRS bins and gout prevalence in males", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)

resulttable_males <- as.data.frame(matrix(ncol=6, nrow=0))
colnames(resulttable_males)<-c("name","OR","LCI","UCI", "P","N" )
males$GoutPRS_bins<-factor(males$GoutPRS_bins, levels =c("[4.73,5.02)","[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)",  "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]"))
test<-glm(gout_V0~GoutPRS_bins+AGE,data=males) 

for(b in seq(1:11)){
    resulttable_males[b,]<-results(test=test,i=b)
}
resulttable_males[1,]<-c("GoutPRS_bins[4.73,5.02)","1","1","1",NA, NA)
resulttable_males[,1]<-c("[4.73,5.02)","[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)",  "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]", "Age")
resulttable_males$OR<-as.numeric(resulttable_males$OR)
resulttable_males$UCI<-as.numeric(resulttable_males$UCI)
resulttable_males$LCI<-as.numeric(resulttable_males$LCI)
resulttable_males$P<-as.numeric(resulttable_males$P)

resulttable_males$sig<-NA
resulttable_males$sig[which(resulttable_males$P<0.05)]="*"
resulttable_males$sig[which(resulttable_males$P<0.01)]="**"
resulttable_males$sig[which(resulttable_males$P<0.001)]="***"

resulttable_males$name <- factor(resulttable_males$name, levels = c("[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[4.73,5.02)", "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]", "Age"))

library(ggrepel)
ggplot(data=resulttable_males, aes(x=name, y=OR)) +
  geom_bar(stat="identity", color="black") +
  labs(title = "Males", y= "Odds Ratio (Log scale)", x = "Risk Factor") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + 
  scale_y_log10() + 
  geom_errorbar(aes(ymin=LCI, ymax=UCI), width=.2, position=position_dodge(.9)) +
  geom_text_repel(label = resulttable_males$sig, nudge_y = 0.015, aes(segment.alpha=0))

#fix significance
#library(svglite)
#ggsave("~/Desktop/GoutCohort_barchart.svg")

resulttable_males %>%  datatable(., caption = "Logisitic regression Gout ~ GRS bins in males ", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)
```

# females

```{r}
females<-pheno2[which(pheno2$SEX==0),]
# breaksgout2 <- seq(from = range(females$GoutPRS, na.rm = T)[1], to = range(females$GoutPRS, na.rm = T)[2], by = (abs(range(females$GoutPRS, na.rm = T)[1]) + range(females$GoutPRS, na.rm = T)[2])/10)
# bucketing values into bins
# females<-females %>% mutate(GoutPRS_bins = cut(females$GoutPRS, breaks=breaksgout2, include.lowest=TRUE, right=FALSE))
females$GoutPRS_bins<-as.character(females$GoutPRS_bins)

females$GoutPRS_bins[which(females$GoutPRS_bins=="[3.56,3.85)"|females$GoutPRS_bins=="[3.85,4.14)")]="[3.56,4.14)"
females$GoutPRS_bins[which(females$GoutPRS_bins=="[5.9,6.2)"|females$GoutPRS_bins=="[6.2,6.49]")]="[5.9,6.2)"

females$GoutPRS_bins<-as.factor(females$GoutPRS_bins)
females$GoutPRS_bins<-factor(females$GoutPRS_bins, levels = c("[3.56,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[4.73,5.02)", "[5.02,5.32)", "[5.32,5.61)",  "[5.61,5.9)",   "[5.9,6.2)"))

sumfilegout_females<-as.data.frame(table(females$GoutPRS_bins))
sumfilegout_females$controls<-table(females$GoutPRS_bins, females$gout_V0)[1:8]
sumfilegout_females$gout<-table(females$GoutPRS_bins, females$gout_V0)[9:16]
sumfilegout_females$prevalencegout<- round(x = (sumfilegout_females$gout/sumfilegout_females$Freq)*100, digits = 3)

f<-ggplot(sumfilegout_females, aes(Var1,Freq))+geom_col()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20)))
f
a<-sumfilegout_males %>% mutate(Var1_number = as.numeric(Var1)) %>% ggplot( aes(Var1_number,prevalencegout))+geom_line()+
  xlab("Gout risk score bin")+
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + scale_x_continuous(labels = as.character(sumfilegout$Var1), breaks = 1:nlevels(sumfilegout$Var1))
a

sumfilegout_females %>%  datatable(., caption = "GoutPRS bins and gout prevalence in females", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)

resulttable_females <- as.data.frame(matrix(ncol=6, nrow=0))
colnames(resulttable_females)<-c("name","OR","LCI","UCI", "P","N" )
females$GoutPRS_bins<-factor(females$GoutPRS_bins, levels =c("[4.73,5.02)","[3.56,4.14)", "[4.14,4.44)", "[4.44,4.73)",  "[5.02,5.32)", "[5.32,5.61)",  "[5.61,5.9)",   "[5.9,6.2)" ))
test<-glm(gout_V0~GoutPRS_bins+AGE,data=females) 

for(b in seq(1:9)){
    resulttable_females[b,]<-results(test=test,i=b)
}
resulttable_females[1,]<-c("GoutPRS_bins[4.73,5.02)","1","1","1",NA, NA)
resulttable_females[,1]<-c("[4.73,5.02)", "[3.56,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[5.02,5.32)", "[5.32,5.61)",  "[5.61,5.9)",   "[5.9,6.2)", "Age")
resulttable_females$OR<-as.numeric(resulttable_females$OR)
resulttable_females$UCI<-as.numeric(resulttable_females$UCI)
resulttable_females$LCI<-as.numeric(resulttable_females$LCI)
resulttable_females$P<-as.numeric(resulttable_females$P)

resulttable_females$sig<-NA
resulttable_females$sig[which(resulttable_females$P<0.05)]="*"
resulttable_females$sig[which(resulttable_females$P<0.01)]="**"
resulttable_females$sig[which(resulttable_females$P<0.001)]="***"

resulttable_females$name <- factor(resulttable_females$name, levels = c("[3.56,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[4.73,5.02)", "[5.02,5.32)", "[5.32,5.61)",  "[5.61,5.9)",   "[5.9,6.2)", "Age"))

library(ggrepel)
ggplot(data=resulttable_females, aes(x=name, y=OR)) +
  geom_bar(stat="identity", color="black") +
  labs(title = "Females", y= "Odds Ratio (Log scale)", x = "Risk Factor") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + 
  scale_y_log10() + 
  geom_errorbar(aes(ymin=LCI, ymax=UCI), width=.2, position=position_dodge(.9)) +
  geom_text_repel(label = resulttable_females$sig, nudge_y = 0.001, aes(segment.alpha=0))

#fix significance
#library(svglite)
#ggsave("~/Desktop/GoutCohort_barchart.svg")

resulttable_females %>%  datatable(., caption = "Logisitic regression Gout ~ GRS bins in females ", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)
```


# Test Nick's Theory 

```{r}
pheno2$A<-NA
pheno2$A[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$A[which(as.character(pheno2$GoutPRS_bins)=="[3.56,3.85)")]=1
pheno2$B<-NA
pheno2$B[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$B[which(as.character(pheno2$GoutPRS_bins)=="[3.85,4.14)")]=1
pheno2$C<-NA
pheno2$C[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$C[which(as.character(pheno2$GoutPRS_bins)=="[4.14,4.44)")]=1
pheno2$D<-NA
pheno2$D[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$D[which(as.character(pheno2$GoutPRS_bins)=="[4.44,4.73)")]=1

pheno2$E<-NA
pheno2$E[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$E[which(as.character(pheno2$GoutPRS_bins)=="[5.02,5.32)")]=1
pheno2$G<-NA
pheno2$G[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$G[which(as.character(pheno2$GoutPRS_bins)=="[5.32,5.61)")]=1
pheno2$H<-NA
pheno2$H[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$H[which(as.character(pheno2$GoutPRS_bins)=="[5.61,5.9)")]=1
pheno2$I<-NA
pheno2$I[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$I[which(as.character(pheno2$GoutPRS_bins)=="[5.9,6.2)")]=1
pheno2$J<-NA
pheno2$J[which(as.character(pheno2$GoutPRS_bins)=="[4.73,5.02)")]=0
pheno2$J[which(as.character(pheno2$GoutPRS_bins)=="[6.2,6.49]")]=1


testlist<-c("A","B","C","D","E","G","H","I","J")

resulttable <- as.data.frame(matrix(ncol=6, nrow=0))
colnames(resulttable)<-c("name","OR","LCI","UCI", "P","N" )

for (bin in testlist){
resulttable[bin,]<-results(test=glm(gout_V0~pheno2[,bin]+AGE+SEX,data=pheno2))
}

resulttable$OR<-as.numeric(resulttable$OR)
resulttable$UCI<-as.numeric(resulttable$UCI)
resulttable$LCI<-as.numeric(resulttable$LCI)
resulttable$P<-as.numeric(resulttable$P)

resulttable$sig<-NA
resulttable$sig[which(resulttable$P<0.05)]="*"
resulttable$sig[which(resulttable$P<0.01)]="**"
resulttable$sig[which(resulttable$P<0.001)]="***"

resulttable$name <- c("[3.56,3.85)", "[3.85,4.14)", "[4.14,4.44)", "[4.44,4.73)", "[5.02,5.32)" ,"[5.32,5.61)" , "[5.61,5.9)" ,  "[5.9,6.2)" , "[6.2,6.49]")

ggplot(data=resulttable, aes(x=name, y=OR)) +
  geom_bar(stat="identity", color="black") +
  labs(title = "All", y= "Odds Ratio (Log scale)", x = "Risk Factor") +
  theme(plot.title = element_text(size=14, face="bold", hjust = 0.5), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", axis.title.x = element_text(margin = margin(t = 20))) + 
  scale_y_log10() + 
  geom_errorbar(aes(ymin=LCI, ymax=UCI), width=.2, position=position_dodge(.9)) +
  geom_text_repel(label = resulttable$sig, nudge_y = 0.015, aes(segment.alpha=0))

#fix significance
#library(svglite)
#ggsave("~/Desktop/GoutCohort_barchart.svg")

resulttable %>%  datatable(., caption = "Logisitic regression Gout ~ GRS bins ", escape = F, options = list(pageLength = 30, autoWidth = TRUE) , rownames = F)


```


